---
title: "Exploring Appropriate Models for Aflatoxin Values"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
date: "2025-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(readxl)
library(MASS)
library(ggplot2)
```

# Prepare data

```{r}
dat <- read_xlsx("data/Aflatoxin Testing Results 2024-2025_MHC.xlsx")

# Clean column names
dat <- dat %>% 
  rename(Date = `Sample Date`, ticket_num = `Ticket #`, aflatoxin_ppb = `Aflatoxins, ppb`)
```

# Histogram of full data


```{r}
ggplot(dat, aes(x = aflatoxin_ppb)) +
  geom_histogram(aes(y = after_stat(density)), breaks = seq(from = -1, to = 25, by = 1), closed ="right")
```

It appears that there are a lot of 0s! The remaining non-zero values follow some sort of right-skewed distribution.

# The "hurdle" model

General formula (I think)

Let $f_{y!=0}$ by a valid pdf that does not include $y=0$. 

Then,

$$
Pr(Y=y) =
\begin{cases}
p & y = 0 \\
(1-p)f_{y!=0}(y) & y > 0
\end{cases}
$$


# LogNormal model (for positive values)

Extract positive values

```{r}
dat_pos <- dat %>% 
  filter(aflatoxin_ppb > 0)
```


```{r}
# fit the model
fit.lognormal <- fitdistr(dat_pos$aflatoxin_ppb, densfun = "lognormal")
fit.lognormal

# Create table with probability density values for lognormal distribution 
# with parameters from fit.lognormal
xvals <- seq(1e-4, 25, by = 0.1)
lnrom_vals <- dlnorm(x = xvals, 
                     meanlog = fit.lognormal$estimate[1],
       sdlog = fit.lognormal$estimate[2])
lognormal.tb <- data.frame(x = xvals, prob_density = lnrom_vals)

# Plot it!
ggplot(dat_pos, aes(x = aflatoxin_ppb)) +
  geom_histogram(aes(y = after_stat(density)))+
  geom_line(data = lognormal.tb, aes(x = x, y = prob_density), col = "cornflowerblue", linewidth = 1.5)+
  ggtitle("Fit of LogNormal model\n(on positive values only)")
# summary(fit.lognormal)

AIC(fit.lognormal)

```

# Gamma model (for positive values)

```{r}
fit.gamma <- fitdistr(dat_pos$aflatoxin_ppb, densfun = "gamma")
fit.gamma
 #    shape         rate   
 #  1.35631237   0.42181142 
 # (0.13135500) (0.04922118)

# Create table with probability density values for gamma distribution 
# with parameters from fit.gamma
gamma_vals <- dgamma(x = xvals, 
        shape = fit.gamma$estimate[1],
       rate = fit.gamma$estimate[2])
gamma.tb <- data.frame(x = xvals, prob_density = gamma_vals)

# Plot it!
ggplot(dat_pos, aes(x = aflatoxin_ppb)) +
  geom_histogram(aes(y = after_stat(density)))+
  geom_line(data = gamma.tb, aes(x = x, y = prob_density), col = "cornflowerblue", linewidth = 1.5)+
  ggtitle("Fit of Gamma model\n(on positive values only)")
#summary(fit.gamma)

AIC(fit.gamma) # 749.329
```


Gamma models appears to have a slightly better fit both visually and based on the AIC (AIC for Gamma model =749, AIC for LogNormal = 777).

There are other right-skewed distributions that we could try as well!

# Gamma-only model?

What if we fit a gamma model to all of the data? Obs there would still be 0 probability at 0, but maybe we can fit a gamma that would be so highly skewed that it would still work?

```{r}
# First try fitting on raw data with zeros -> I doubt that this would work, but let's try it
# fit.gamma.full <- fitdistr(dat$aflatoxin_ppb, densfun = "gamma")
# error

# Set 0 values to a very small positive number
# Seems like data values go u pto 3 decimal places 
# So let's true 1e-4
dat$aflatoxin_ppb_no_zero <- ifelse(dat$aflatoxin_ppb > 0, dat$aflatoxin_ppb, 1e-4)
fit.gamma.v2 <- fitdistr(dat$aflatoxin_ppb_no_zero, densfun = "gamma")
fit.gamma.v2
# Warning that NaNs produced...

# How doe sthe fit look?
xvals <- seq(1e-4, 25, by = 0.1)
gamma_vals <- dgamma(x = xvals, 
        shape = fit.gamma.v2$estimate[1],
       rate = fit.gamma.v2$estimate[2])
gamma.tb <- data.frame(x = xvals, prob_density = gamma_vals)

# Plot it!
ggplot(dat, aes(x = aflatoxin_ppb_no_zero)) +
  geom_histogram(aes(y = after_stat(density)))+
  geom_line(data = gamma.tb, aes(x = x, y = prob_density), col = "cornflowerblue", linewidth = 1.5)+
  ggtitle("Fit of Gamma model\n(zeros set to 1e-4)")+
  ylim(0, 30)
#summary(fit.gamma)

# Wow, this is HIGHLY right-skewed!!!
# I'm goign to stick wit hthe hurdle model for now
```

# Show hist full data, gamma on positives

```{r}
xvals <- seq(1e-4, 25, by = 0.1)
gamma_vals <- dgamma(x = xvals, 
        shape = fit.gamma$estimate[1],
       rate = fit.gamma$estimate[2])
gamma.tb <- data.frame(x = xvals, prob_density = gamma_vals)

ggplot(dat, aes(x = aflatoxin_ppb)) +
  geom_histogram(mapping = aes(y = after_stat(density)), closed = "right")+
  geom_line(data = gamma.tb, aes(x = x, y = prob_density), col = "cornflowerblue", linewidth = 1.5)+
  ggtitle("Fit of Gamma model\n(on positive values only)")

ggplot(dat, aes(x = aflatoxin_ppb)) +
  geom_histogram(aes(y = after_stat(density)), breaks = seq(from = -1, to = 25, by = 1), closed ="right")+
  geom_line(data = gamma.tb, aes(x = x, y = prob_density), col = "cornflowerblue", linewidth = 1.5)+
  ggtitle("Fit of Gamma model\n(on positive values only)")
#summary(fit.gamma)
```

