---
title: "Exploring Appropriate Models for Aflatoxin Values"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
date: "2025-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(readxl)
library(MASS)
library(ggplot2)
```

# Prepare data

```{r}
dat <- read_xlsx("data/Aflatoxin Testing Results 2024-2025_MHC.xlsx")

# Clean column names
dat <- dat %>% 
  rename(Date = `Sample Date`, ticket_num = `Ticket #`, aflatoxin_ppb = `Aflatoxins, ppb`)
```

# Histogram of full data


```{r}
ggplot(dat, aes(x = aflatoxin_ppb)) +
  geom_histogram(aes(y = after_stat(density)))
```

It appears that there are a lot of 0s! The remaining non-zero values follow some sort of right-skewed distribution.

# The "hurdle" model

General formula (I think)

Let $f_{y!=0}$ by a valid pdf that does not include $y=0$. 

Then,

$$
Pr(Y=y) =
\begin{cases}
p & y = 0 \\
(1-p)f_{y!=0}(y) & y > 0
\end{cases}
$$


# LogNormal model (for positive values)

Extract positive values

```{r}
dat_pos <- dat %>% 
  filter(aflatoxin_ppb > 0)
```


```{r}
# fit the model
fit.lognormal <- fitdistr(dat_pos$aflatoxin_ppb, densfun = "lognormal")
fit.lognormal

# Create table with probability density values for lognormal distribution 
# with parameters from fit.lognormal
xvals <- seq(1e-4, 25, by = 0.1)
lnrom_vals <- dlnorm(x = xvals, 
                     meanlog = fit.lognormal$estimate[1],
       sdlog = fit.lognormal$estimate[2])
lognormal.tb <- data.frame(x = xvals, prob_density = lnrom_vals)

# Plot it!
ggplot(dat_pos, aes(x = aflatoxin_ppb)) +
  geom_histogram(aes(y = after_stat(density)))+
  geom_line(data = lognormal.tb, aes(x = x, y = prob_density), col = "cornflowerblue", linewidth = 1.5)+
  ggtitle("Fit of LogNormal model\n(on positive values only)")
# summary(fit.lognormal)

AIC(fit.lognormal)

```

# Gamma model (for positive values)

```{r}
fit.gamma <- fitdistr(dat_pos$aflatoxin_ppb, densfun = "gamma")
fit.gamma

# Create table with probability density values for gamma distribution 
# with parameters from fit.gamma
gamma_vals <- dgamma(x = xvals, 
        shape = fit.gamma$estimate[1],
       rate = fit.gamma$estimate[2])
gamma.tb <- data.frame(x = xvals, prob_density = gamma_vals)

# Plot it!
ggplot(dat_pos, aes(x = aflatoxin_ppb)) +
  geom_histogram(aes(y = after_stat(density)))+
  geom_line(data = gamma.tb, aes(x = x, y = prob_density), col = "cornflowerblue", linewidth = 1.5)+
  ggtitle("Fit of Gamma model\n(on positive values only)")
#summary(fit.gamma)

AIC(fit.gamma) # 749.329
```


Gamma models appears to have a slightly better fit both visually and based on the AIC (AIC for Gamma model =749, AIC for LogNormal = 777).

There are other right-skewed distributions that we could try as well!