---
title: "Rough Draft: Simulation part 2"
output: html_document
date: "2025-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(ggplot2)
```

# Goal

Specifically, we want to simulate:

* Take a scoop from every individual railcar in a train (assume homogeneity within each railcar)
* Composite the scoops into groups of X and calculate the average aflatoxin level in the composites.
* Decide to accept or reject the trainload based on whether any composite has more than 20 ppb



# Define functions for simulation

Function to simulate trainload with fixed parameters

```{r}
# Function to simulate one trainload
simulate_trainload <- function(nrailcars_per_train, 
                               prob_of_zero,
                               shape_param, 
                               scale_param) {
  
  # Step one: determine if the value is 0 or not
  # (returns list of 0s and 1s)
  value_is_nonzero_indicator <- rbinom(n = nrailcars_per_train, 
                                       size = 1,
                                       p = 1-prob_of_zero)
  
  # Step two: draw from gamma dist
  gamma_component <- rgamma(nrailcars_per_train, 
                            shape = shape_param, 
                            scale = scale_param)  
  
  # Step three: Multiply together to get a single vector
  observed <- value_is_nonzero_indicator*gamma_component
  
  return(observed)
}
```

Function to loop through grid of parameter values and simulate several trainloads for each

```{r}
# Define combinations of parameter values to simulation
param.tb <- expand.grid(prob_of_zero = seq(0, 0.05, by = 0.05), # 2 
                        shape_param = seq(1, 4, by = 1), # 4
                        scale_param = seq(1, 5, by = 1), # 5
                        num_railcars_per_composite = c(5, 10, 20, 40))

# Function to simulate 
run_simulation <- function(n_sim_per_param, param.tb, 
                           nrailcars_per_train = 200, 
                           regulatory_threshold = 20) {
  
  # Progress bar setup
  pb <- txtProgressBar(min = 0, max = nrow(param.tb), style = 3)
  
  # Initialize table to save output at each simulation
  sim.tb <- data.frame()
  
  # Loop over parameter combinations
  for (i in 1:nrow(param.tb)) {
    
    parameters <- param.tb[i,]

    for (j in 1:n_sim_per_param) {
      
      trainload_sim_ppb <- simulate_trainload(nrailcars_per_train, 
                                              parameters$prob_of_zero,
                                              parameters$shape_param, 
                                              parameters$scale_param)

      # True safe proportion in the full trainload
      train_true_safe_prop <- mean(trainload_sim_ppb < regulatory_threshold)
      
      # Simulate getting composite samples
      num_composites <- nrailcars_per_train / parameters$num_railcars_per_composite
      trainload_sim_tb <- data.frame(values = trainload_sim_ppb,
                                     composite_id = rep(1:num_composites, 
                                                        each = parameters$num_railcars_per_composite ))
      facility_sim_tb <- trainload_sim_tb %>% 
        group_by(composite_id) %>% 
        reframe(afla_in_composite = mean(values))

      # Save information on the simulation
      add.tb <- data.frame(parameter_id = i,
                           rep_id = j,
                           
                           # True trainload data
                           train_true_safe_prop = train_true_safe_prop,
                           
                           # Facility data
                           prop_safe_composites = mean(facility_sim_tb$afla_in_composite < regulatory_threshold),
                           composites_avg = mean(facility_sim_tb$afla_in_composite),
                           composites_sd = sd(facility_sim_tb$afla_in_composite))
      sim.tb <- rbind(sim.tb, add.tb)
      
    }
    
    setTxtProgressBar(pb, i)  # Update progress bar
  }
  close(pb)
  return(sim.tb)
}
```

# Run simulation 

```{r eval = FALSE}
set.seed(454)
res2 <- run_simulation(n_sim_per_param = 500, param.tb)
save(res2, file = "sim_data/simulation2_parameter_grid_Apr9_2025.RData")
```

Clean results

```{r}
load("sim_data/simulation2_parameter_grid_Apr9_2025.RData")

# Add labels
res2 <- res2 %>% 
  mutate(train_is_safe = train_true_safe_prop == 1)
param.tb$parameter_id <- 1:nrow(param.tb)
res2 <- res2 %>% 
  left_join(param.tb, by = "parameter_id")
res2 <- res2 %>%
  mutate(trainload_pass = prop_safe_composites == 1)

# Calculate power, etc
summary2.tb <- res2 %>% 
  group_by(parameter_id, prob_of_zero, shape_param, scale_param, num_railcars_per_composite, train_is_safe) %>% 
  reframe(total_trains = n(), 
          num_pass = sum(trainload_pass),
          num_fail = sum(!trainload_pass),
          frac_pass = sum(trainload_pass)/n(),
          frac_pass_moe = 1.96*sqrt(frac_pass*(1-frac_pass)/total_trains),
          frac_pass_ci_lb = frac_pass - frac_pass_moe,
          frac_pass_ci_ub = frac_pass + frac_pass_moe)
```

# Plots/results for rough draft

## Figure 4

```{r}
# Add identifier to link parameter combinations that are the same except for 'num railcars per composite'
summary2.tb <- summary2.tb %>% 
  mutate(parameter_id_sans_num_railcars_per_composite = paste(prob_of_zero, shape_param, scale_param, sep = "_"),
         parameter_id_sans_scale = paste(prob_of_zero, shape_param, sep = "_"))

# Prepare data toplot
plotdat <- summary2.tb %>% 
  filter(!train_is_safe, scale_param > 1, prob_of_zero == 0.05)

# Denote scale more clearly
plotdat <- plotdat %>% 
  mutate(scale_plot = paste("scale:",scale_param))
plotdat$shape = as.factor(plotdat$shape_param)

# facet by scale, color by shape
ggplot(plotdat, 
       aes(x = num_railcars_per_composite, y = frac_pass, color = shape))+
  geom_point()+
  geom_line(aes(group = parameter_id_sans_num_railcars_per_composite))+
  facet_wrap(vars(scale_plot))+
  ylab("Type 1 Error Rate")+
  xlab("Number of Railcars per Composite")+
  theme_bw()+
  theme(text = element_text(size = 15))

ggsave(filename = "figs/type1_error_rate_vs_num_railcars_per_composite_facet_scale.png", width = 6, height = 4)
```

See number of unsafe trains per parameter combo, when scale >= 3

```{r}
summary2.tb %>% 
  filter(scale_param >= 3, !train_is_safe) %>% 
  select(total_trains) %>% 
  summary
 # total_trains  
 # Min.   : 92.0  
 # 1st Qu.:471.2  
 # Median :500.0  
 # Mean   :448.6  
 # 3rd Qu.:500.0  
 # Max.   :500.0
```

Cool, so at minimum, there were at least 92 unsafe trainloads for parameter combinations with scale >= 3.

## Figure 5

Show some trainloads

```{r}
example.dat <- sapply(2:5, simulate_trainload, nrailcars_per_train = 200,
                             prob_of_zero = 0.05, shape_param = 4)
colnames(example.dat) <- paste0("scale: ",2:5)
example.dat <- as.data.frame(example.dat)

example.dat <- example.dat %>% 
  pivot_longer(paste0("scale: ",2:5), values_to = "ppb")

ggplot(example.dat, aes(x = ppb)) +
  geom_histogram()+
  geom_vline(xintercept = 20, color = "blue")+
  facet_wrap(vars(name))+
  xlab("Aflatoxin level (ppb)")+
  ylab("Number of railcars")+
  theme_bw()+
  theme(text = element_text(size = 15))

ggsave(filename = "figs/example_trainload_dist_prob_0.05_shape_4_scale_facet.png", width = 6, height = 4)
```


## Confirm 100% of safe trains accepted

```{r}
summary2.tb %>% 
  filter(train_is_safe) %>% 
  select(frac_pass) %>% 
  summary

```

Yep, 100% of simulated trainloads that were safe passed.

I guess this makes sense - if the trainload truly is safe, then there is no way that any composite sample will be greater than 20.


# (old) Exploratory plots/results

```{r}
summary2.tb <- summary2.tb %>% 
  mutate(parameter_id_sans_num_railcars_per_composite = paste(prob_of_zero, shape_param, scale_param, sep = "_"),
         parameter_id_sans_scale = paste(prob_of_zero, shape_param, sep = "_"))

ggplot(summary2.tb, aes(x = num_railcars_per_composite, y = frac_pass, color = train_is_safe))+
  geom_point()+
  geom_line(aes(group = parameter_id_sans_num_railcars_per_composite))+
  facet_wrap(vars(train_is_safe))
# Once again, 100% of the safe trains pass (i.e., producers risk or Type 2 error rate is always 0)

# Focus on type 1 error rate

# Which param(s) seem to be controlling the results the most?
# prob of zero - no
ggplot(summary2.tb[summary2.tb$train_is_safe == FALSE,], 
       aes(x = num_railcars_per_composite, y = frac_pass, color = as.factor(shape_param)))+
  geom_point()+
  geom_line(aes(group = parameter_id_sans_num_railcars_per_composite))+
  facet_wrap(vars(train_is_safe))

ggplot(summary2.tb[summary2.tb$train_is_safe == FALSE,], 
       aes(x = num_railcars_per_composite, y = frac_pass, color = as.factor(scale_param)))+
  geom_point()+
  geom_line(aes(group = parameter_id_sans_num_railcars_per_composite))+
  facet_wrap(vars(train_is_safe))



# Possibly some kidn of ineraction effect between the shape and scale param....
# let's color by shape, x-axis is scale, then facet by num_railcars_per_composite
summary.unsafe <- summary2.tb[summary2.tb$train_is_safe == FALSE,]

ggplot(summary.unsafe, aes(x = scale_param, y = frac_pass, color = as.factor(shape_param)))+
  geom_point()+
  geom_line(aes(group = parameter_id_sans_scale))+
  geom_errorbar(aes(ymin = frac_pass_ci_lb, ymax = frac_pass_ci_ub))+
  facet_wrap(vars(num_railcars_per_composite))


ggplot(summary.unsafe, aes(x = scale_param, y = total_trains, color = as.factor(shape_param)))+
  geom_point()+
    geom_line(aes(group = parameter_id_sans_scale))+
  facet_wrap(vars(num_railcars_per_composite))
```

```{r}
ggplot(summary.unsafe, 
       aes(x = num_railcars_per_composite, y = frac_pass, color = as.factor(scale_param)))+
  geom_point()+
  geom_line(aes(group = parameter_id_sans_num_railcars_per_composite))+
    geom_errorbar(aes(ymin = frac_pass_ci_lb, ymax = frac_pass_ci_ub))+
  facet_wrap(vars(shape_param))

summary2.tb %>% 
  filter(scale_param == 1, train_is_safe == FALSE) %>% 
  select(prob_of_zero, shape_param, train_is_safe, total_trains, num_pass)
#  When scale_param == 1, there were max of 2 unsafe trains
# So we can ignore this case

# facet by scale, color by shape
ggplot(summary.unsafe[summary.unsafe$scale_param > 1,], 
       aes(x = num_railcars_per_composite, y = frac_pass, color = as.factor(shape_param)))+
  geom_point()+
  geom_line(aes(group = parameter_id_sans_num_railcars_per_composite))+
    geom_errorbar(aes(ymin = frac_pass_ci_lb, ymax = frac_pass_ci_ub))+
  facet_wrap(vars(scale_param))

```

Why is it that when the scale param == 2, 100% of unsafe trains are accepted? That seems odd to me... 
Let's look at a distribution

```{r}
dat_unsafe <- simulate_trainload(200, 0, shape_param = 4, scale_param = 2)
hist(dat_unsafe)
```

I guess the idea is that when the variability is relatively low, then if any railcars are high, it's unlikely that any individual composite will show as being unsafe, even if you have relatively few samples per composite! That's a cool takeaway!!

